---
title: "Visualization of Simulation Results"
output: word_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=8}
# this script performs visualization of the simulation results

library(tidyverse)
library(ggpubr)
library(viridis)

sim_results <- readRDS("sim_results.rds")
sim_results <- sim_results %>% mutate(admin = ifelse(admin=="pill", "oral", admin))
```

# Simulation Setup

There are 4 different scenarios that are simulated. They differ in: 

  * the decision rule (liberal vs. conservative, see next paragraph for a discussion) and 
  * the recruitment speed (20 vs. 30 patients per month). 

All other settings are identical for all simulations. Particularly, 

  * the sample size is 90 patients with option to stop for futility after 45 patients,
  * all trials use concurrent controls only,
  * the assumptions for TRD and PRD are the same, 
  * effect sizes are the same for the three domains,
  * expressed in standardized mean difference between baseline and 6-weeks MADRS score, effect sizes according to Cohen are: 
      + d = 0.00
      + d = 0.22
      + d = 0.35
      + d = 0.50
      + the correlation between the two measuremnts is assumed to be rho = 0.214
      + a new treatment is randomly assigned one of the four effect sizes with equal probability for each.  
  * there cannot be more than 8 treatments in the platform in parallel,
  * each month, if there are less than 8 active treatments in the platform, one new treatment might be added with probabilities 0.2 for oral, 0.1 for nasal and 0.1 for IV. 
  * allocation ratio treatment:control within each domain is 1:sqrt(K), where K is the total number of treatments - however allocation to control is never less than 35% regardless of the number of treatments,
  * the platform does not add any more compounds at time points later than 60 months since platform start,
  * after this time point, the platform ends once all compounds have made a decision. 

# Decisions

Terminology: In the following, "conservative" refers to a setting with more strict decision rules, i.e., two-sided p-values of p<0.05 compared to control to claim success of a compound and p<0.4 to allow continuation to second stage, and "liberal" refers to a setting with less strict decision rules, i.e., two-sided p-values of p<0.1 compared to control to claim success of a compound and p<0.5 to allow continuation to second stage.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=8}
scenario_labeller <- labeller(
  `cohens_d` = c(`0` = "Cohen's d = 0", `0.22` = "Cohen's d = 0.22", `0.35` = "Cohen's d = 0.35", `0.5` = "Cohen's d = 0.50"))

sim_results %>% filter(patients_per_timepoint==30 & pvalues=="0.4,0.05") %>% 
  filter(treatment_ID != "Control") %>% 
  mutate(decisions_TRD = factor(decisions_TRD, levels=c("success", "stopped early", "failure")),
         cohens_d = factor(round(cohens_d_TRD,2))) %>% 
  group_by(admin, cohens_d, decisions_TRD, N, .drop=FALSE) %>% summarise(n=n()) %>%
  group_by(admin, cohens_d, N) %>% mutate(percentage = 100 * n / sum(n)) %>% 
  {ggplot(., aes(admin, percentage, color=decisions_TRD)) + geom_point(size=2, alpha = 1, position=position_dodge(width=0.5)) + 
      geom_text(data=subset(., decisions_TRD=="success"), aes(admin, percentage, label=round(percentage,2)), size=3, vjust=1.5, hjust=0.7) +
      scale_y_continuous(limits=c(0,100), breaks = seq(0,100,by=20), minor_breaks = seq(0,100,by=10)) + 
      xlab("Way of administration") + 
      scale_color_viridis_d(begin = 0.1, end = 0.9, name="Decision") +
      ylab("Percentage") + #ggtitle(title) +
      geom_hline(data = data.frame(y = c(5, 80, 80, 80), cohens_d = as.factor(c(0,0.22,0.35,0.5))), aes(yintercept=y), linetype="dotted") +
      #geom_errorbar(aes(Prior, y=Proportion, ymin=lower, ymax=upper), size=0.5, position=position_dodge(width=0.5)) + 
      facet_grid( ~ cohens_d, labeller = scenario_labeller) +
      theme_bw() + ggtitle("Conservative decision rules:\nPercentage of each decision for various true effect sizes and under different sample sizes") +
      theme(legend.position = "bottom")}

sim_results %>% filter(patients_per_timepoint==30 & pvalues=="0.5,0.1")  %>% 
  filter(treatment_ID != "Control") %>% 
  mutate(decisions_TRD = factor(decisions_TRD, levels=c("success", "stopped early", "failure")),
         cohens_d = factor(round(cohens_d_TRD,2))) %>% 
  group_by(admin, cohens_d, decisions_TRD, N, .drop=FALSE) %>% summarise(n=n()) %>%
  group_by(admin, cohens_d, N) %>% mutate(percentage = 100 * n / sum(n)) %>% 
  {ggplot(., aes(admin, percentage, color=decisions_TRD)) + geom_point(size=2, alpha = 1, position=position_dodge(width=0.5)) + 
      geom_text(data=subset(., decisions_TRD=="success"), aes(admin, percentage, label=round(percentage,2)), size=3, vjust=1.5, hjust=0.7) +
      scale_y_continuous(limits=c(0,100), breaks = seq(0,100,by=20), minor_breaks = seq(0,100,by=10)) + 
      xlab("Way of administration") + 
      scale_color_viridis_d(begin = 0.1, end = 0.9, name="Decision") +
      ylab("Percentage") + #ggtitle(title) +
      geom_hline(data = data.frame(y = c(5, 80, 80, 80), cohens_d = as.factor(c(0,0.22,0.35,0.5))), aes(yintercept=y), linetype="dotted") +
      #geom_errorbar(aes(Prior, y=Proportion, ymin=lower, ymax=upper), size=0.5, position=position_dodge(width=0.5)) + 
      facet_grid( ~ cohens_d, labeller = scenario_labeller) +
      theme_bw() + ggtitle("Liberal decision rules:\nPercentage of each decision for various true effect sizes and under different sample sizes") +
      theme(legend.position = "bottom")}


ggsave("decisions_TRD.tiff", device = "tiff", width=9, height=4)

```

For the other statistical endpoints, the decision threshold is not overly influential. Its only impact is that slightly more trials are stopped early due to futility when the decision threshold is more conservative. Therefore, for the remaining analysis, the simulations with liberal decision criterion are considered. Results are mostly very similar to the simulation with conservative decision criterion.  

```{r, echo=FALSE, message=FALSE, warning=FALSE}
sim_results <- sim_results %>% filter(pvalues=="0.5,0.1")
```


# Sample sizes

## Total number of patients per domain

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=8}

### aggregated sample sizes 

N_per_admin <- sim_results %>%  
  group_by(nsim, patients_per_timepoint, N, admin) %>% summarise(n_TRD = sum(n_TRD), n_PRD = sum(n_PRD))

N_per_admin %>% 
  pivot_longer(., cols=starts_with("n_"), names_to="pop", names_prefix="n_", values_to="n") %>%
  filter(pop=="TRD") %>%
  ggplot(.) + geom_violin(aes(x=admin, y=n, fill=admin)) +
  facet_grid(~patients_per_timepoint, labeller=labeller(
    `patients_per_timepoint` = c(`20` = "Recruitment per month: 20 patients", `30` = "Recruitment per month: 30 patients"))) +
  theme_bw() + ggtitle("Total number of patients per domain") + ylab("") + xlab("Domain") + theme(legend.position="none")
ggsave("N_per_domain.tiff", device = "tiff", width=9, height=4)

```


## Total number of patients per population

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=8}
N_per_pop <- sim_results %>% group_by(nsim, patients_per_timepoint, N) %>% summarise(n_TRD = sum(n_TRD), n_PRD = sum(n_PRD))
N_per_pop %>% 
  pivot_longer(., cols=starts_with("n_"), names_to="pop", names_prefix="n_", values_to="n") %>%
  ggplot(.) + geom_violin(aes(x=as.factor(pop), y=n, fill=pop)) +
  facet_grid(~patients_per_timepoint, labeller=labeller(
    `patients_per_timepoint` = c(`20` = "Recruitment per month: 20 patients", `30` = "Recruitment per month: 30 patients"))) +
  theme_bw() + ggtitle("Total number of patients per population") + ylab("") + xlab("") + theme(legend.position="none")
ggsave("N_per_pop.tiff", device = "tiff", width=9, height=4)

```



## Total number of patients in platform

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=8}
N_per_platform <- sim_results %>% group_by(nsim, patients_per_timepoint, N) %>% summarise(n_total = sum(n_TRD+n_PRD))

N_per_platform %>% 
  ggplot(.) + geom_violin(aes(x="", y=n_total), fill="grey") +
  facet_grid(~patients_per_timepoint, labeller=labeller(
    `patients_per_timepoint` = c(`20` = "Recruitment per month: 20 patients", `30` = "Recruitment per month: 30 patients"))) +
  theme_bw() + ggtitle("Total number of patients per platform") + ylab("") + xlab("") + theme(legend.position="none")
ggsave("N_per_platform.tiff", device = "tiff", width=9, height=4)

```


## Sample size of control arms per population

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=8}

sim_results %>% 
  filter(treatment_ID=="Control") %>% 
  ggplot(.) + geom_violin(aes(x=admin, y=n_TRD, fill=admin)) +
  facet_grid(~patients_per_timepoint, labeller=labeller(
    `patients_per_timepoint` = c(`20` = "Recruitment per month: 20 patients", `30` = "Recruitment per month: 30 patients"))) +
  theme_bw() + ylab("Sample Size") + xlab("Domain")  + theme(legend.position="none")
ggsave("sample_size_control_population.tiff", device = "tiff", width=9, height=3)

```



## Sample size of treatment arms for final decision (over-recruiting)

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=8}

sim_results %>% 
  filter(treatment_ID!="Control") %>% 
  ggplot(.) + geom_bar(aes(x=n_treatment_TRD, fill=admin), position = position_dodge2()) +
  facet_grid(~patients_per_timepoint, labeller=labeller(`patients_per_timepoint` = c(`20` = "Recruitment per month: 20 patients", `30` = "Recruitment per month: 30 patients"))) +
  theme_bw() + ylab("Count") + xlab("Sample size of treatment arm for final decision") + labs(fill="Domain") + theme(legend.position = "right")
ggsave("sample_size_control_population.tiff", device = "tiff", width=9, height=3) 

```


# Size of control arms used for final decision (excl. stopping for futility)


```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=8}

sim_results %>% 
  filter(decisions_TRD %in% c("success","failure")) %>% 
  ggplot(.) + geom_bar(aes(x=n_control_comparators_TRD, fill=admin), position = position_dodge2()) +
  facet_grid(~patients_per_timepoint, labeller=labeller(
    `patients_per_timepoint` = c(`20` = "Recruitment per month: 20 patients", `30` = "Recruitment per month: 30 patients")
  )) +
  theme_bw() + xlab("Size of control arm used for final decision") + ylab("Count") + labs(fill="Domain") + theme(legend.position = "right")

ggsave("size_control_final_analysis.tiff", device = "tiff", width=9, height=3)

```



# Number of arms

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=8}
number_of_arms_per_admin <- sim_results %>% group_by(nsim, patients_per_timepoint, N, admin) %>% count()
number_of_arms_per_admin %>% group_by(patients_per_timepoint, N, admin, "n"=as.factor(n), .drop=F) %>% 
  summarise(counts=n()) %>%
  mutate(counts=counts/max(sim_results$nsim)*100) %>%
  ggplot(.) + 
  geom_bar(aes(x=as.factor(n), y=counts, fill=admin), stat="identity", position = "dodge", width=0.7) + 
  facet_grid(~patients_per_timepoint, labeller=labeller(
    `patients_per_timepoint` = c(`20` = "Recruitment per month: 20 patients", `30` = "Recruitment per month: 30 patients"))) + coord_flip() +
  theme_bw() + ggtitle(paste0("Number of tested compounds per platform")) + ylab("Percentage of simulations") + xlab("Number of compounds") + 
  scale_fill_viridis_d(end=0.9, name="Domain")
ggsave("treatments_per_admin.tiff", device = "tiff", width=9, height=4)
```


```{r, include=FALSE, eval=FALSE}
# Number of arms in platform (TRD and PRD are not counted separately)
number_of_arms_per_platform <- sim_results %>% group_by(nsim, scenarioID) %>% count()
```



# Estimation of Cohen's d

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=8}
scenario_labeller <- labeller(
  `cohens_d` = c(`0` = "Cohen's d = 0", `0.22` = "Cohen's d = 0.22", `0.35` = "Cohen's d = 0.35", `0.5` = "Cohen's d = 0.50"),
  `admin` = c(`IV`="IV", `nasal`="nasal", `oral`="oral")
)
sim_results %>% filter(patients_per_timepoint==30) %>%
  filter(treatment_ID != "Control") %>% 
  mutate(cohens_d = factor(round(cohens_d_TRD,2))) %>%
  ggplot(.) + geom_violin(aes(x=as.factor(N), y=cohens_d_TRD_est), fill="grey") +
  geom_hline(data = data.frame(y = c(0, 0.22, 0.35, 0.5), cohens_d = as.factor(c(0,0.22,0.35, 0.5))), aes(yintercept=y), linetype="dotted") +
  facet_grid(admin~cohens_d, labeller = scenario_labeller) +
  theme_bw() + ylab("Estimated Cohen's d") + ggtitle("Estimated Cohen's d") + xlab("Sample Size") + scale_fill_discrete(name="N") + theme(legend.position = "none")
ggsave("estimated_cohens_d.tiff", device = "tiff", width=9, height=3)

```
The estimation of the effect size is subject to variance, which is slightly reduced by the increase of the sample size from 80 to 100.


```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=8}

scenario_labeller <- labeller(
  `cohens_d` = c(`0` = "Cohen's d = 0", `0.22` = "Cohen's d = 0.22", `0.35` = "Cohen's d = 0.35", `0.5` = "Cohen's d = 0.50")
)
sim_results %>% filter(patients_per_timepoint==30) %>%
  filter(treatment_ID != "Control") %>% 
  mutate(cohens_d = factor(round(cohens_d_TRD,2))) %>%
  group_by(admin, N, cohens_d) %>% summarise(est_cohens_d = mean(cohens_d_TRD_est)) %>%
  ggplot(.) + geom_point(aes(x=admin, y=est_cohens_d, fill=admin)) +
  geom_hline(data = data.frame(y = c(0, 0.22, 0.35, 0.5), cohens_d = as.factor(c(0,0.22,0.35, 0.5))), aes(yintercept=y), linetype="dotted") +
  facet_grid(~cohens_d, labeller = scenario_labeller) +
  theme_bw() + ylab("Mean of estimated Cohen's d") + ggtitle("Bias in estimation of Cohen's d") + xlab("Domain") + theme(legend.position="none")
ggsave("estimated_cohens_d.tiff", device = "tiff", width=9, height=3)

```

Bias is a systematic over- or underestimation of an estimator. The true value as specified in the simulations is indicated by the dotted lines. The average estimation is indicated by the dot. Naturally, when there is early stopping for futility, bias is introduced. The bias is smaller for large effect sizes, because the trials rarely stops. 


# Duration

## Average duration per arm

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=8}
sim_results %>% filter(treatment_ID != "Control") %>% 
  group_by(cohens_d=round(cohens_d_TRD,2), patients_per_timepoint, admin) %>%
  summarise(duration=mean(duration_TRD)) %>%
  # when wrapping the whole ggplot call in {...} curly braces, you must use the . dot pronoun for the data argument in ggplot(., ...). Then you can call back that object using the . pronoun anywhere in the call.
  # https://stackoverflow.com/questions/45088454/how-do-i-access-the-data-frame-that-has-been-passed-to-ggplot
  {ggplot(.) + geom_segment(aes(x=0, xend=duration, y=admin, yend=admin), size=2) +
      facet_grid(cohens_d~patients_per_timepoint, labeller=labeller(
        `cohens_d` = c(`0` = "Cohen's d = 0", `0.22` = "Cohen's d = 0.22", `0.35` = "Cohen's d = 0.35", `0.5` = "Cohen's d = 0.50"),
        `patients_per_timepoint` = c(`20` = "Recruitment per month: 20 patients", `30` = "Recruitment per month: 30 patients"))) + 
      theme_bw() + xlab("Duration (in months)") + ylab("Domain") + 
      scale_x_continuous(breaks=seq(0,max(.$duration),by=10))}

```


```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=8, fig.width=8}
sim_results %>% filter(treatment_ID != "Control") %>% 
  group_by(cohens_d=round(cohens_d_TRD,2), patients_per_timepoint, admin) %>%
  ggplot(.) + geom_violin(aes(x=duration_TRD, y=admin, fill=admin)) +
  facet_grid(cohens_d~patients_per_timepoint, labeller=labeller(
        `cohens_d` = c(`0` = "Cohen's d = 0", `0.22` = "Cohen's d = 0.22", `0.35` = "Cohen's d = 0.35", `0.5` = "Cohen's d = 0.50"),
        `patients_per_timepoint` = c(`20` = "Recruitment per month: 20 patients", `30` = "Recruitment per month: 30 patients"))) + 
  theme_bw() + xlab("Duration (in months)") + ylab("Domain") + theme(legend.position = "none")
```


## Average duration per arm with interval

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.width=8}
sim_results %>% group_by(N, patients_per_timepoint, admin) %>%
  summarise(duration_mean=mean(duration_TRD),
            duration_q10=quantile(duration_TRD,probs=0.1),
            duration_q90=quantile(duration_TRD,probs=0.9)) %>%
  pivot_longer(., cols=starts_with("duration"), names_prefix="duration_", names_to="quantity", values_to="duration") %>%
  mutate(quantity = factor(quantity, levels = c("q10", "mean", "q90"))) %>%
  # when wrapping the whole ggplot call in {...} curly braces, you must use the . dot pronoun for the data argument in ggplot(., ...). Then you can call back that object using the . pronoun anywhere in the call.
  # https://stackoverflow.com/questions/45088454/how-do-i-access-the-data-frame-that-has-been-passed-to-ggplot
  {ggplot(.) + geom_linerange(aes(xmin=0, xmax=duration, y=admin, color=quantity), position = position_dodge(.5), size=2) +
      facet_grid(~patients_per_timepoint, labeller=labeller(
        `patients_per_timepoint` = c(`20` = "Recruitment per month: 20 patients", `30` = "Recruitment per month: 30 patients"))) + theme_bw() + xlab("Duration (in months)") + ylab("Domain") + 
      scale_x_continuous(breaks=seq(0,max(.$duration),by=10)) + 
      scale_color_discrete(name="Quantity", labels=c("10% quantile", "Mean duration", "90% quantile")) +
      theme(legend.position = "right", legend.direction="vertical")}

```


## Duration per arm for selected simulations

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=8}
set.seed(11)
sim_results %>% filter(nsim==sample(unique(.$nsim), 1) & scenarioID==4) %>%
  select(admin, treatment_ID, contains("timestamp")) %>%
  pivot_longer(., cols=c(contains("first_timestamp"), contains("last_timestamp")), names_to=c("timestamp", "pop"),names_pattern = "(.+)_(.+)") %>%
  pivot_wider(., names_from="timestamp", values_from="value") %>%
  ggplot(.) + geom_segment(aes(x=first_timestamp, xend=last_timestamp, y=treatment_ID, yend=treatment_ID), size=2) +
  facet_wrap(pop~admin, scales="free_y", labeller = label_wrap_gen(multi_line=FALSE)) + theme_bw() + 
  xlab("Time point") + ylab("Treatment Arm") + ggtitle("Duration per arm for one selected simulation")
sim_results %>% filter(nsim==sample(unique(.$nsim), 1) & scenarioID==4) %>%
  select(admin, treatment_ID, contains("timestamp")) %>%
  pivot_longer(., cols=c(contains("first_timestamp"), contains("last_timestamp")), names_to=c("timestamp", "pop"),names_pattern = "(.+)_(.+)") %>%
  pivot_wider(., names_from="timestamp", values_from="value") %>%
  ggplot(.) + geom_segment(aes(x=first_timestamp, xend=last_timestamp, y=treatment_ID, yend=treatment_ID), size=2) +
  facet_wrap(pop~admin, scales="free_y", labeller = label_wrap_gen(multi_line=FALSE)) + theme_bw() + 
  xlab("Time point") + ylab("Treatment Arm") + ggtitle("Duration per arm for one selected simulation")
sim_results %>% filter(nsim==sample(unique(.$nsim), 1) & scenarioID==4) %>%
  select(admin, treatment_ID, contains("timestamp")) %>%
  pivot_longer(., cols=c(contains("first_timestamp"), contains("last_timestamp")), names_to=c("timestamp", "pop"),names_pattern = "(.+)_(.+)") %>%
  pivot_wider(., names_from="timestamp", values_from="value") %>%
  ggplot(.) + geom_segment(aes(x=first_timestamp, xend=last_timestamp, y=treatment_ID, yend=treatment_ID), size=2) +
  facet_wrap(pop~admin, scales="free_y", labeller = label_wrap_gen(multi_line=FALSE)) + theme_bw() + 
  xlab("Time point") + ylab("Treatment Arm") + ggtitle("Duration per arm for one selected simulation")
```

These are just 3 randomly chosen examples. One could look at multiple such illustrations to get an impression of how the platform can potentially proceed. 

